A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-22 doc/paper.tex
Case study rewording. Fixes #75, fixes #76.

	\section{Related Work} Given humanity's affinity for visual information,
	visualizations have long been prized for their ability to communicate
	information quickly and efficient \cite{baeza1999modern}; the visualization of
	spreadsheets, therefore, has offered much room for research. Several previous
	tools work within a spreadsheet, seeking to clarify the connections between the
	visible values in cells, the formulae that produce them, and the dependencies
	between cells that inform them. Igarashi and colleague's fluid visualizations
	\cite{igarashi1998fluid}, for example, do this by imposing lines, color, and
	animation over the spreadsheet to visually explain these tangled connections.
	Likewise, Clermont explored ways of grouping cells by color and border that
	were related through similar functions, neighbors, or references
	\cite{clermont2003scalable}, and Hipfl extended this work by incorporating the
	layouts and labels of spreadsheets \cite{hipfl2008using}. Other tools focus on
	creating visualizations external to the sheets. Hermans and colleagues address
	a spreadsheet programmer's information by making dataflow diagrams from
	individual sheets through the tools GyroSAT \cite{hermans2011supporting} and
	Breviz \cite{hermans2011breviz}. \toolname is among the latter class, creating
	standalone visualizations about spreadsheets. In contrast to Hermans' work,
	though, our work focuses on representing functions and how they comprise
	complex formula, rather than exploring the data structures of any particular
	spreadsheet.  \par

@@

	For both sides of the equal sign, the operator has certain types of arguments
	which predominate over the others: on the left side is most often a reference
	to another cell; on the right, a string or number literal, which makes sense
	for the case of confirming a value in another cell before assigning this one.
	Furthermore, if the concept is difficult to imagine is practice, a tooltip
	accompanies each function node in the tree, providing a concrete example of a
	function that uses this structure and where it can be found. If this single
	instance is not enough, then the user also has the option to double-click the
	individual function node, which will open up a new tab with a table of many
	more examples from dataset.

@@

		%			\begin{itemize}
		%
		%				\item [1] \textbf{Provide an interactive environment for free
		% exploration.}
		%				There are a lot of data in some of these collections, and the user should
		% be
		%				able to pick for themselves which to explore while ignoring the rest.
		% Furthermore,
		%				this exploration should not be designed around answering one specific
		% question
		%				but rather be open-ended.
		%
		%				\item [!1] \textbf{Do not make judgments about good and bad design.}
		% Though
		%				users may certainly approach the tool with the intention of finding
		% particular
		%				formulae, \toolname will not prioritize any on metrics of good or bad
		%				practice. Rather, it will present them agnostically, referring only to
		%				frequency.
		%
		%				\item [2] \textbf{Emphasize the quantitative patterns in formula
		%					construction.} \toolname should address the questions of how often the
		%				spreadsheet programmers used a certain function and where they used it.
		% In
		%				this way, it must show precise metrics, such as frequency of use and
		% depth
		% of
		%				function nesting, of the dataset it conveys.
		%
		%				\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		%				especially, we anticipated some particularly dense trees. Though
		%				everything should be accessible in accordance with Goal 1, we should
		% design
		% in
		%				a way that reduces noise.
		%
		%				\item [3] \textbf{Promote a qualitative understanding of the patterns.}
		%				\toolname should supplement its observations with concrete instances of
		%				relevant formulae from the corpus. Ideally, it should even direct users
		% to
		% the
		%				exact cell in the spreadsheets where the formula was used,
		% contextualizing
		% the
		%				functions.
		%
		%				\item [!3] \textbf{Do not attempt to explain formulae.} Though \toolname
		% tries
		%				to foster understanding by linking pattern to example, it should not
		% provide
		%				a precise description of what a complex formula accomplishes. The tool's
		% user
		%				must infer this on their own.
		%
		%			\end{itemize}

@@

	\section{Case Study} The obvious question to pose to any visualization is this:
	Why does it matter that we see this way? In other words, it is not enough to
	prove that something can be visualized; instead we must also demonstrate what
	we can be gained from any particular visualization. As such, we raised a few
	tasks in the introduction in which we thought the tool could help: identifying
	places for API improvement, detecting bad smells or common programmer
	misunderstanding, and guiding spreadsheet education. To demonstrate the tool's
	applicability, then, we sought out places in the tree which best exemplify
	these respective concerns.

	\subsection{Bad Smells} Fowler's description of code smells
	\cite{fowler2009refactoring} underscored an important point of code quality:
	between perfect code and bug-crippled spaghetti,  there is a spectrum of code
	designs which, by themselves, are not faulty but nevertheless suggest problems
	and vulnerabilities in code design. Code smells, as we call them, soon
	commanded attention and earned wide-ranging applicability, eventually making it
	to the art of spreadsheet design.

	Since then, spreadsheet researchers have taken the concept, if not the specific
	smells themselves, and applied them to spreadsheet programming
	\cite{hermans2012detecting} \cite{cunha2012towards} \cite{jansen2015code}
	\cite{asavametha2012detecting} . While some of the smells lie beyond the scope
	of the tool, such as those characterizing inter-worksheet connections, some
	have structures which create distinctive features within the tree, leading to
	easy detection through visualization: \par

	\begin{itemize} \item \textit{Long Method}: Also called "Multiple Operations"
		in \cite{hermans2012detecting}, this refers to situations where numerous
		functions and operators are used all within the same cell, creating a formula
		that tends to extend beyond a compact viewing space and makes it difficult to
		understand. In the visualization, since each nested and combined function
		creates yet another level of depth within the tree, these constructions can
		result in long, horizontal chains of nodes, as seen in figure
		\ref{"fig:longsum"}. \item \textit{Long Parameter List}: [tall, vertical towers
		of blue nodes] \item \textit{Conditional Complexity}: [IF nodes within IF
		nodes] \end{itemize}

>>>>

	\section{Related Work} \label{related-work} Given humanity's affinity for visual information,
	visualizations have long been prized for their ability to communicate
	information quickly and efficient \cite{baeza1999modern}; the visualization of
	spreadsheets, therefore, has offered much room for research. Several previous
	tools work within a spreadsheet, seeking to clarify the connections between the
	visible values in cells, the formulae that produce them, and the dependencies
	between cells that inform them. Igarashi and colleague's fluid visualizations
	\cite{igarashi1998fluid}, for example, do this by imposing lines, color, and
	animation over the spreadsheet to visually explain these tangled connections.
	Likewise, Clermont explored ways of grouping cells by color and border that
	were related through similar functions, neighbors, or references
	\cite{clermont2003scalable}, and Hipfl extended this work by incorporating the
	layouts and labels of spreadsheets \cite{hipfl2008using}. Other tools focus on
	creating visualizations external to the sheets. Hermans and colleagues address
	a spreadsheet programmer's information by making dataflow diagrams from
	individual sheets through the tools GyroSAT \cite{hermans2011supporting} and
	Breviz \cite{hermans2011breviz}. \toolname is among the latter class, creating
	standalone visualizations about spreadsheets. In contrast to Hermans' work,
	though, our work focuses on representing functions and how they comprise
	complex formula, rather than exploring the data structures of any particular
	spreadsheet.  \par

@@

	For both sides of the equal sign, the operator has certain types of arguments
	which predominate over the others: on the left side is most often a reference
	to another cell; on the right, a string or number literal, which makes sense
	for the case of confirming a value in another cell before assigning this one.
	Furthermore, a tooltip accompanies each function node in the tree, providing a
	concrete example of a function that uses this structure and where it can be
	found. If this single instance is not enough, then the user also has the option
	to double-click the individual function node, which will open up a new tab with
	a table of many more examples from dataset.

@@

		%Can I incorporate these again?
		%			\begin{itemize}
		%
		%				\item [1] \textbf{Provide an interactive environment for free
		% exploration.}
		%				There are a lot of data in some of these collections, and the user should
		% be
		%				able to pick for themselves which to explore while ignoring the rest.
		% Furthermore,
		%				this exploration should not be designed around answering one specific
		% question
		%				but rather be open-ended.
		%
		%				\item [!1] \textbf{Do not make judgments about good and bad design.}
		% Though
		%				users may certainly approach the tool with the intention of finding
		% particular
		%				formulae, \toolname will not prioritize any on metrics of good or bad
		%				practice. Rather, it will present them agnostically, referring only to
		%				frequency.
		%
		%				\item [2] \textbf{Emphasize the quantitative patterns in formula
		%					construction.} \toolname should address the questions of how often the
		%				spreadsheet programmers used a certain function and where they used it.
		% In
		%				this way, it must show precise metrics, such as frequency of use and
		% depth
		% of
		%				function nesting, of the dataset it conveys.
		%
		%				\item [!2] \textbf{Do not overwhelm with numbers.} For popular functions
		%				especially, we anticipated some particularly dense trees. Though
		%				everything should be accessible in accordance with Goal 1, we should
		% design
		% in
		%				a way that reduces noise.
		%
		%				\item [3] \textbf{Promote a qualitative understanding of the patterns.}
		%				\toolname should supplement its observations with concrete instances of
		%				relevant formulae from the corpus. Ideally, it should even direct users
		% to
		% the
		%				exact cell in the spreadsheets where the formula was used,
		% contextualizing
		% the
		%				functions.
		%
		%				\item [!3] \textbf{Do not attempt to explain formulae.} Though \toolname
		% tries
		%				to foster understanding by linking pattern to example, it should not
		% provide
		%				a precise description of what a complex formula accomplishes. The tool's
		% user
		%				must infer this on their own.
		%
		%			\end{itemize}

@@

	\section{Case Study} The obvious question to pose to any visualization is this:
	What do the images tell us that the text could not? It is not enough to
	prove that something can be visualized; we must also demonstrate what
	we can be gained from any particular visualization. As such, we raised a few
	tasks in the introduction in which we thought the tool could help: identifying
	places for API improvement, detecting bad smells, and guiding spreadsheet education. To demonstrate the tool's
	applicability, we sought out places in the tree which best exemplify
	these tasks.

	\subsection{Bad Smells} Fowler's description of code smells
	\cite{fowler2009refactoring} underscored an important point of code quality:
	between perfect code and bug-crippled spaghetti,  there is a spectrum of code
	designs which, by themselves, are not faulty but nevertheless suggest vulnerabilities in code design. 	
	Since then, researchers have applied these concepts to spreadsheet programming, as 
	discussed in section \ref{related-work}. While some of the smells lie beyond the scope
	of the tool, such as those characterizing inter-worksheet connections, others
	have structures which create distinctive features within the tree, leading to
	easy detection through our visualization: \par

		\item \textit{Long Method}: Also called "Multiple Operations"
		in \cite{hermans2012detecting}, this refers to situations where numerous
		functions and operators are used within a single formula. As a result, the formula
		tends to be prohibitively complex. In the visualization, since each nested and combined function
		creates yet another level of depth within the tree, these constructions can
		result in long, horizontal chains of nodes, as seen in figure
		\ref{"fig:longsum"}.

		\item \textit{Long Parameter List}: [tall, vertical towers
		of blue nodes]

		\item \textit{Conditional Complexity}: [IF nodes within IF
		nodes] \end{itemize}
